Robust Estimation, Prediction and Control with \\ Linear Dynamics and Generic Costs

### 2020-02-03 main.tex
Update

% Custom packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{blkarray}
\usepackage{mathtools}
\input{mathdef.tex}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

@@

To address this issue, we turn to the framework of \emph{robust} decision-making: instead of merely considering a point estimate of the dynamics, we build an entire \emph{confidence region} that contains the true dynamics with high probability, and optimise the decisions so as to maximise the worst-case outcome with respect to this uncertainty. Minimax problem are hard to solve, especially since the reward function $R$ can be arbitrary. In particular, this keeps us from employing traditional tools from the robust synthesis literature, such as $\cH_\infty$ \todo{(ref needed)} control or system-level synthesis \todo{(ref needed)}. Since we are optimising a generic reward function, we turn to more general sequential optimisation algorithms. The continuous state space prevents us from using dynamic programming techniques, so we consider tree-based planning instead. Though some work address continuous actions \todo{ref needed}, we resort to a first approximation and discretise the decision space by adopting a hierarchical control architecture. We  formulate the problem high-level decision of selecting a low-level controller among a finite basis \todo{(ref from temporal abstraction)}. For instance, this could amount to choosing a short-term goal to follow, or a skill to execute such as a driving maneuver. The passage from continuous to discrete optimization induces a suboptimality, but it can be mitigated by diversifying the controller basis. These low-level controllers are not learnt, which also constitutes a limitation \todo{realistic in industrial setting?}. However, tree-based planning algorithms are designed to optimise a singled known generative model, and not a whole confidence region. Hence, we adapt them to the robust framework by following a two step procedure: we first predict of the set of states that can be reached by following a control sequence, under all possible dynamics. This reachability set is used to provide a conservative pessimistic reward to the planning algorithm.

@@

In \autoref{sec:estimation}, having observed history of transitions $\cD_{[N]} = \{(x_n, u_n, y_n)\}_{n\in[N]}$, and given a confidence level $\delta\in(0, 1)$, our goal is to find a confidence region $\cC_{[N],\delta}$ as tight as possible and such that it holds that:
\begin{align}
\probability{A(\theta)\in \cC_{[N],\delta}} \geq 1-\delta
\label{eq:confidence}
\end{align}
Our first contribution is to extend the work of \citet{Abbasi2011} that provide a confidence ellipsoid for the regularised least-square estimator in the case of scalar noise and vector features to our setting of vector noise and matrix features.

@@

At the time step $n=N$, considering the model uncertainty $\cC_{[N],\delta}$ obtained in the model estimation phase, we formulate our robust control objective:
\begin{equation}
\label{eq:robust-control}
\sup_{\bu\in{(\Real^q)}^\Natural} \underbrace{\inf_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bw\in[\underline{\bw},\overline{\bw}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bw))\right]}_{V^r(\bu)}
\end{equation}
where $\bu = (u_{N+1},u_{N+2},\cdots)$ and $\bw = (\omega_N, \omega_{N+1},\cdots)$ are the sequences of future controls and perturbations, respectively.

The rest of the paper will tackle the optimisation of this objective.

@@

Minimax problems such as \eqref{eq:robust-control} are notoriously difficult to solve even in cases such as Linear-Quadratic systems (LQ) where the reward $R$ has a simple form. Here with an arbitrary function $R$, we cannot hope to derive a solution in explicit form. Our second contribution is to propose to solve \eqref{eq:robust-control} numerically by adopting a \emph{predict-then-optimise} scheme. Thus, in \autoref{sec:prediction}, we aim to derive a \emph{set predictor} $X(t)$ for the system \eqref{eq:dynamics}, which takes the information on the observed current state ${x}_N$, a confidence region $\cC_\delta$ containing $A(\theta)$, a planned control sequence $\bu$ and some admissible bounds on the perturbation $[\underline{\omega}(t),\overline{\omega}(t)]$; and produces predictions that verify the inclusion property:
\begin{equation}
 x(t)\in X(t),\quad\forall t\geq t_N.
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.3\textwidth]{img/interval-hull}
    \caption{A few trajectories are sampled from an initial state $x_N$ following controls $\bu=(u_0,\cdots)$ with various parameters $A(\theta)\in\cC_{[N],\delta}$ (in black). At time $t\geq t_N$, any reachable state $x(t)$ (in green) must belong to the candidate set predictor $X(t)$ (in red).}
    \label{fig:interval-hull}
\end{figure}

@@

Since $R$ is arbitrary, solving the optimal -- not to mention the robust -- control objective in the original continuous control space $(\Real^q)^\Natural$ is intractable.
In this work, we choose to approximate the solution by turning the continuous optimisation problem \eqref{eq:robust-control} into a discrete one, by optimising over a finite set of \emph{actions} $\cA$. Each action $a\in\cA$ corresponds to the selection of a low-level controller $\pi_a$, that we take affine:
\[u(t) = \pi_a(x(t)) \eqdef -K_a x(t) + u_a\]

@@

In \autoref{sec:control}, facing an optimisation problem with continuous states and discrete actions, we turn to the literature of tree-based planning. Our third contribution proposes to approximate the true robust objective ${V}^r$ by an easily-evaluable surrogate $\hat{V}^r$ that exploits the set predictor to define a pessimistic reward, used to evaluate sequences of actions in a tree-based planning algorithm. We show that

% \paragraph{Objective}

% We wish to find a sequence of commands $\bu=(u_t)_{t\geq 0}$ that maximises a cumulative objective $V^\bu$:

% \todo[inline]{EL: the prediction is done in continuous time and the control in discrete time. How to handle the change in notations?}

% \begin{equation}
% \max_\bu V^\bu
% \end{equation}
% where
% \begin{equation}
% \label{eq:optimal-control}
% V^\bu = \expectedvalueover{\omega(t)}\left[\sum_{t=0}^\infty \gamma^t r(x(t))\condbar u_t,\,\eqref{eq:dynamics}\right]
% \end{equation}
% and $r:\Real^p\rightarrow[0,1]$ is a bounded reward function.

\section{Related Work}

@@

In the general case, 
 assume that the dynamics $A(\theta)$ enjoy the following additional structure:
\begin{assumption}[Structure]
\label{assumpt:structure}
There exists a known feature tensor $\phi\in \Real^{d \times p \times p}$ such that for all $\theta\in\Theta$:
\begin{equation}
    A(\theta) = A + %\theta^\transp \phi \eqdef A + 
    \sum_{i=1}^d \theta_i\phi_i
\end{equation}
where $\phi_i\in\Real^{p\times p}$ for all $i\in[d]$. For all $n$, we also denote $\Phi_n = [\phi_1 x_n \cdots \phi_d x_n]\in\Real^{p\times d}$.

@@

\begin{proposition}[Regularised solution]
\label{prop:regularized_solution}
The solution to \eqref{eq:regression_min} is:
\begin{equation}
    \label{eq:vector_rls}
    \theta_{N,\lambda} = G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} y_n
\end{equation}
where 
\begin{equation*}
    G_{N, \lambda} = \sum_{n=1}^N \Phi_{n}^\transp\Sigma_p^{-1}\Phi_{n}  + \lambda I_d \in \Real^{d\times d}.
\end{equation*}
\end{proposition}

@@

\begin{theorem}[Confidence ellipsoid, matrix version of \citep{Abbasi2011}]
\label{thm:confidence_ellipsoid}
Under \autoref{assumpt:gaussian-noise}, it holds with probability at least $1-\delta$ that
\begin{align}
    \label{eq:confidence-ellipsoid}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} \leq \beta_N(\delta),
\end{align}
with
\begin{equation}
    \beta_N(\delta)\eqdef \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
     + (\lambda d)^{1/2}S.
\end{equation}
\end{theorem}

@@

A simple solution to \eqref{eq:inclusion-property} is proposed in \citep{efimov:hal-00701643}, where they use matrix interval arithmetics to derive the predictor:
\begin{proposition}
Assuming that \eqref{eq:confidence} is satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & \underline{A}^{+}\underline{x}^{+}(t)-\overline{A}^{+}\underline{x}^{-}(t)-\underline{A}^{-}\overline{x}^{+}(t) +\overline{A}^{-}\overline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\label{eq:IP_direct}\\
\dot{\overline{x}}(t) & = & \overline{A}^{+}\overline{x}^{+}(t)-\underline{A}^{+}\overline{x}^{-}(t)-\overline{A}^{-}\underline{x}^{+}(t)+\underline{A}^{-}\underline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(0)=\underline{x}_{0},\;\overline{x}(0)=\overline{x}_{0},\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:inclusion-property}.
\end{proposition}

@@

% \begin{example*}
% Consider a scalar system:
% \[
% \dot{x}(t)=-\theta(t)x(t)+d(t),\;t\geq0,
% \]
% where $x(t)\in\Real$ with $x(0)\in[\underline{x}_{0},\overline{x}_{0}]=[1.0, 1.1]$, $\theta(t)\in\Pi=[\underline{\theta},\overline{\theta}]=[0.5,1.5]$ and $d(t)\in[\underline{d},\overline{d}]=[-0.1,0.1]$ for all $t\geq0$. The \autoref{fig:predictor_example} shows the state-interval obtained with \eqref{eq:interval-predictor}.

% \begin{figure}
% \begin{centering}
% \includegraphics[width=\linewidth]{img/interval-predictor}
% \par\end{centering}
% \caption{Several trajectories for different values of $\theta\in\Theta$ are shown in blue. The result of prediction by \eqref{eq:interval-predictor} is in red: the predictor is stable and produces tight bounds.}
% \label{fig:predictor_example}
% \end{figure}
% \end{example*}

Limitations: set-based method assumes time-varying uncertainty where we have time-invariant uncertainty: $A(\theta)$ is fixed vs $A(\theta)$ can vary in $\cC_\delta$ at each step. Consequence: conservative overset.

@@

However, under an additional smoothness assumption on $R$, we can bound this gap:
\begin{proposition}
\label{prop:control-error}
If the reward function $R$ is lipschitz, then the approximation error is upper-bounded by:
\begin{equation*}
     V^r(\bu) - \hat{V}^r(\bu) = \cO\left(\frac{1}{\sqrt{N}}\right)
\end{equation*}
\end{proposition}

\begin{algorithm}[tb]
	\caption{Robust Estimation, Prediction and Control}
	\label{alg:full}
	\begin{algorithmic}
		\STATE {\bfseries Input:} confidence level $\delta$
		\FOR{$N = 1,2,3,\dots$} 
		\STATE $\cC_{[N],\delta} \gets$\textsc{Model Estimation}$(\cD_{[N]})$. \eqref{eq:polytope}
		\FOR{planning steps $k\in[K]$}
		\STATE $a_k\gets$\textsc{Pessimistic Planning}$(\underline{R_n}([\underline{x}_{n}, \overline{x}_{n}]))$.  \eqref{eq:opd}
		\FOR{$b\in A$}
		\STATE $[\underline{x}_{n+1}(a_k b), \overline{x}_{n+1}(a_k b)]\gets$ \textsc{Interval Prediction}($\cC_{[N],\delta}$). \eqref{eq:interval-predictor}
		\ENDFOR
		\ENDFOR
		\STATE Execute the first recommended control $u_N$, and add the observed transition $(x_N, u_N, \dot{x}_N)$ to $\cD_{[N+1]}$.
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

@@

Given $N-1$ observations, the \autoref{sec:estimation} provides a polytopic confidence region \eqref{eq:polytope} that contains $A(\theta)$ with probability at least $1-\delta$. Since the dynamics are linear, we can propagate this confidence region to the next observation: $y_{N}$ must belong to the Minkowski sum of a polytope representing model uncertainty $\cP(A_{0} x_N + Bu_N, \Delta A_{1}x_N,\cdots \Delta A_{2^d}x_N)$ and a polytope $\cP(0_p, \underline{\eta}, \overline{\eta})$ bounding the perturbation and measurement noises. \citet{delos2015} provide a way to test this membership in polynomial time using linear programming.

Whenever this membership property is not verified, we can confidently reject the $(A,\phi)$-modelling \autoref{assumpt:structure}. This enables us to consider a rich set of potential features $\left((A^1, \phi^1), \cdots, (A^M, \phi^M)\right)$ rather than relying on a single assumption.

@@

\begin{assumption}[Multi-model ambiguity]
\label{assumpt:multi-model-ambiguity}
Assume that the true dynamics $f$ lies within a finite set of candidate models $f^1, \cdots, f^M$.
\begin{equation}
\exists m\in[M]: \dot{x}(t) = f^M(x(t), u(t)), \forall t\geq 0
\end{equation}
\end{assumption}
We want to adapt our planning algorithm in order to balance these concurrent hypotheses in a robust fashion, i.e. maximise a robust objective with discrete ambiguity:
\begin{equation}
\label{eq:robust-objective-discrete}
V^r = \sup{\bu}\min_{m\in[M]} \sum_{t=1}^\infty \gamma^t R_t^m
\end{equation}
where $R_t^m$ is the reward obtain by following the control sequence $\bu$ up to step $t$ under the dynamics $f^m$.
This objective could be optimised in the same way as \ref{sec:control}, by computing a hull over the $M$ possible trajectories, but this would result in a coarse and lossy approximation. Instead, we can exploit the finite uncertainty structure of \autoref{assumpt:multi-model-ambiguity} to asymptotically recover the true $V^r$ by modifying the \texttt{OPD} planning algorithm in the following way:

@@

Note that this theorem bounds the performance gap with respect to the number of node expansions, but each expansion still requires $M$ time more simulations in the multi-model setting than in the single-model setting. The solution of the robust objective \eqref{eq:robust-objective-discrete} with discrete ambiguity $f\in\{f^m\}_{m\in[M]}$ can be recovered exactly, asymptotically when the planning budget $H$ goes to infinity. This is in contrast to the results obtained \autoref{sec:control} regarding the robust objective \eqref{eq:robust-control} with continuous ambiguity $A(\theta)\in\cC_{[N],\delta}$, for which there is a gap between the asymptotic performance and the true robust value, as discussed in \autoref{prop:control-error}.

@@

% \paragraph{Obstacle avoidance with unknown friction}
% We first consider a simple illustrative example: the control of a $2D$-

@@

@@

\subsection{Proof of \autoref{prop:control-error}}

\begin{proof}
 By the inclusion property \eqref{eq:inclusion-property}, for any sequence of controls $\bu$, dynamics $A(\theta)\in C_{[N],\delta}$ and perturbations $\underline{\bw} \leq \bw \leq \overline{\bw}$, we have that $\underline{x}_n \leq x_n \leq \overline{x}_n$, which implies that $R(x_n) \leq \max_{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}  R(x)$.

 Thus, since we assume that $R$ is $L$-lipschitz,
 \begin{align}
     V^r(\bu) - \hat{V}^r(\bu) &\leq \sum_{n=N+1}^\infty \gamma^n \underset{{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}}{(\max - \min)} R(x)\\
     &\leq \sum_{n=N+1}^\infty \gamma^n L \left\|\underline{x}_n(\bu) - \overline{x}_n(\bu)\right\|
 \end{align}

>>>>

% Custom packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{blkarray}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{siunitx}
\input{mathdef.tex}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

@@

To address this issue, we turn to the framework of \emph{robust} decision-making: instead of merely considering a point estimate of the dynamics, we build an entire \emph{confidence region} that contains the true dynamics with high probability, and optimise the decisions so as to maximise the worst-case outcome with respect to this uncertainty. Minimax problem are hard to solve, especially since the reward function $R$ can be arbitrary. In particular, this keeps us from employing traditional tools from the robust synthesis literature, such as $\cH_\infty$ \todo{(ref needed)} control or system-level synthesis \todo{(ref needed)}. Since we are optimising a generic reward function, we turn to more general sequential optimisation algorithms. The continuous state space prevents us from using dynamic programming techniques, so we consider tree-based planning instead. Though some work address continuous actions \todo{ref needed}, we resort to a first approximation and discretise the decision space by adopting a hierarchical control architecture. We  formulate the problem high-level decision of selecting a low-level controller among a finite basis \todo{(ref from temporal abstraction)}. For instance, this could amount to choosing a short-term goal to follow, or a skill to execute such as a driving maneuver. The passage from continuous to discrete optimization induces a suboptimality, but it can be mitigated by diversifying the controller basis. These low-level controllers are not learnt, which also constitutes a limitation \todo{(argue it is realistic in industrial setting?)}. However, tree-based planning algorithms are only designed for a single known generative model instead of an entire confidence region for the system dynamics. Hence, we adapt them to the robust framework by following a two-steps procedure: we first predict a set of states that can be reached under a planned control sequence with respect to all possible dynamics. This reachability set is used to provide a conservative pessimistic reward to the planning algorithm.

@@

In \autoref{sec:estimation}, having observed history of transitions $\cD_{[N]} = \{(x_n, y_n)\}_{n\in[N]}$, and given a confidence level $\delta\in(0, 1)$, our goal is to find a confidence region $\cC_{[N],\delta}$ as tight as possible and such that it holds that:
\begin{align}
\probability{A(\theta)\in \cC_{[N],\delta}} \geq 1-\delta
\label{eq:confidence}
\end{align}
Our first contribution is to extend the work of \citet{Abbasi2011} that provide a confidence ellipsoid for the regularised least-square estimator to our setting of vector observations and matrix features, instead of scalar observations and vector features.

@@

At the time step $n=N$, considering the confidence region $\cC_{[N],\delta}$ obtained in the model estimation phase, we formulate our robust control objective:
\begin{equation}
\label{eq:robust-control}
\sup_{\bu\in{(\Real^q)}^\Natural} \underbrace{\inf_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bw\in[\underline{\bw},\overline{\bw}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bw))\right]}_{V^r(\bu)}
\end{equation}
where $\bu = (u_{N+1},u_{N+2},\dots)$ and $\bw = (\omega_N, \omega_{N+1},\dots)$ are the sequences of future controls and perturbations, respectively. The rest of the paper will tackle the optimisation of this objective.

@@

Minimax problems such as \eqref{eq:robust-control} are notoriously difficult to solve even in cases such as Linear-Quadratic systems (LQ) where the reward $R$ has a simple form. Here with an arbitrary function $R$, we cannot hope to derive a solution in explicit form. Our second contribution is to propose to solve \eqref{eq:robust-control} numerically by adopting a \emph{predict-then-optimise} scheme. Thus, in \autoref{sec:prediction}, we aim to derive a \emph{set predictor} $X(t)$ for the system \eqref{eq:dynamics}, which takes the information on the observed current state ${x}_N$, a confidence region $\cC_\delta$ verifying \eqref{eq:confidence}, a planned control sequence $\bu$ and some admissible bounds on the perturbation $[\underline{\omega}(t),\overline{\omega}(t)]$; and produces predictions of sets $X(t)$ that verify the inclusion property:
\begin{equation}
 x(t)\in X(t),\quad\forall t\geq t_N.
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{img/epc-friction}
    \caption{\textbf{Left}: the state prediction procedure. \textbf{Right}: the model estimation procedure.}
    \label{fig:interval-hull}
\end{figure}

@@

Since $R$ is generic, potentially non-smooth and non-convex, solving the optimal -- not to mention the robust -- control objective in the original continuous control space $(\Real^q)^\Natural$ is intractable.
In this work, we choose to approximate the solution by turning the continuous optimisation problem \eqref{eq:robust-control} into a discrete one, by optimising over a finite set of \emph{actions} $\cA$. Each action $a\in\cA$ corresponds to the selection of a low-level controller $\pi_a$, that we take affine:
\[u(t) = \pi_a(x(t)) \eqdef -K_a x(t) + u_a\]

@@

In \autoref{sec:control}, facing an optimisation problem with continuous states and discrete actions, we turn to the literature of tree-based planning. Our third contribution proposes to approximate the true robust objective ${V}^r$ by an easily-evaluable surrogate $\hat{V}^r$ that exploits the set predictor to define a pessimistic reward, used to evaluate sequences of actions in a tree-based planning algorithm.

The \autoref{alg:full} shows the full integration of the three procedures of estimation, prediction and control.

\paragraph{Model selection} In \autoref{sec:multi-model}, we question our modelling assumption and relax them to a more flexible setting.

\paragraph{Experiments} We illustrate the \autoref{alg:full} in a simple 2D example, before moving to a more realistic simulation of an autonomous driving application.

\begin{algorithm}[tb]
	\caption{Robust Estimation, Prediction and Control}
	\label{alg:full}
	\begin{algorithmic}
		\STATE {\bfseries Input:} confidence level $\delta$
		\FOR{$N = 1,2,3,\dots$} 
		\STATE $\cC_{[N],\delta} \gets$\textsc{Model Estimation}$(\cD_{[N]})$. \eqref{eq:polytope}
		\FOR{planning steps $k\in[K]$}
		\FOR{$a\in A$}
		\STATE $[\underline{x}_{n+1}, \overline{x}_{n+1}]\gets$ \textsc{Interval Prediction}(
		\STATE $\qquad\qquad\qquad\qquad\cC_{[N],\delta}, a_k$). \eqref{eq:interval-predictor}
		\ENDFOR
		\STATE $a_{k+1}$ $\gets$\textsc{Pessimistic Planning}$($
		\STATE$\qquad\qquad\underline{R_{n+1}}([\underline{x}_{n+1}, \overline{x}_{n+1}]))$.  \eqref{eq:opd}
		\ENDFOR
		\STATE Execute the first recommended control $u_N$, and add the observed transition $(x_N, y_N)$ to $\cD_{[N+1]}$.
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\subsection{Related Work}

@@

In the general case, 
 assume that the dynamics $A(\theta)$ enjoy the following additional structure:
\begin{assumption}[Structure]
\label{assumpt:structure}
There exists a known feature tensor $\phi\in \Real^{d \times p \times p}$ such that for all $\theta\in\Theta$:
\begin{equation}
    A(\theta) = A + %\theta^\transp \phi \eqdef A + 
    \sum_{i=1}^d \theta_i\phi_i
\end{equation}
where $\phi_i\in\Real^{p\times p}$ for all $i\in[d]$. For all $n$, we also denote $\Phi_n = [\phi_1 x_n \dots \phi_d x_n]\in\Real^{p\times d}$.

@@

\begin{proposition}[Regularised solution]
\label{prop:regularized_solution}
The solution to \eqref{eq:regression_min} is:
\begin{equation}
    \label{eq:vector_rls}
    \theta_{N,\lambda} = G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} y_n
\end{equation}
where 
\begin{equation}
\label{eq:g_n_lambda}
    G_{N, \lambda} = \sum_{n=1}^N \Phi_{n}^\transp\Sigma_p^{-1}\Phi_{n}  + \lambda I_d \in \Real^{d\times d}.
\end{equation}
\end{proposition}

@@

\begin{theorem}[Confidence ellipsoid, matrix version of \citep{Abbasi2011}]
\label{thm:confidence_ellipsoid}
Under \autoref{assumpt:gaussian-noise}, it holds with probability at least $1-\delta$ that
\begin{align}
    \label{eq:confidence-ellipsoid}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} \leq \beta_N(\delta),
\end{align}
with
\begin{equation}
	\label{eq:beta_n}
    \beta_N(\delta)\eqdef \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
     + (\lambda d)^{1/2}S.
\end{equation}
\end{theorem}

@@

A simple solution to \eqref{eq:inclusion-property} is proposed in \citep{efimov:hal-00701643}, where, given bounds $\underline{A}\leq A(\theta)\leq\overline{A}$ on $A(\theta)$ from $\cC_{[N],\delta}$ they use matrix interval arithmetics to derive the predictor:
\begin{proposition}
Assuming that \eqref{eq:confidence} is satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & \underline{A}^{+}\underline{x}^{+}(t)-\overline{A}^{+}\underline{x}^{-}(t)-\underline{A}^{-}\overline{x}^{+}(t) +\overline{A}^{-}\overline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\label{eq:predictor-naive}\\
\dot{\overline{x}}(t) & = & \overline{A}^{+}\overline{x}^{+}(t)-\underline{A}^{+}\overline{x}^{-}(t)-\overline{A}^{-}\underline{x}^{+}(t)+\underline{A}^{-}\underline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(t_N)=\underline{x}_{t_N},\;\overline{x}(t_N)=\overline{x}_{t_N},\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:inclusion-property}.
\end{proposition}

@@

Limitations: set-based method assumes time-varying uncertainty where we have time-invariant uncertainty: $A(\theta)$ is fixed vs $A(\theta)$ can vary in $\cC_\delta$ at each step. Consequence: conservative overset.
\begin{figure}[tp]
	\centering
	\includegraphics[width=\linewidth]{img/interval-predictor}
	\caption{Comparison of \eqref{eq:predictor-naive} to \eqref{eq:interval-predictor} for a simple system $\dot{x}(t)=-\theta x(t)+d(t)$}
	\label{fig:predictor_example}
\end{figure}

@@

However, under an additional smoothness assumption on $R$, we can bound this gap:
\begin{theorem}
\label{thm:control-error}
If the reward function $R$ is lipschitz, then the approximation error of \autoref{alg:full} is upper-bounded by:
\begin{equation*}
     V^r(\bu) - \hat{V}^r(\bu) = \cO\left(\frac{1}{\sqrt{N}}\right)
\end{equation*}
\end{theorem}

@@

Given $N-1$ observations, the \autoref{sec:estimation} provides a polytopic confidence region \eqref{eq:polytope} that contains $A(\theta)$ with probability at least $1-\delta$. Since the dynamics are linear, we can propagate this confidence region to the next observation: $y_{N}$ must belong to the Minkowski sum of a polytope representing model uncertainty $\cP(A_{0} x_N + Bu_N, \Delta A_{1}x_N,\dots, \Delta A_{2^d}x_N)$ and a polytope $\cP(0_p, \underline{\eta}, \overline{\eta})$ bounding the perturbation and measurement noises. \citet{delos2015} provide a way to test this membership in polynomial time using linear programming.

Whenever this membership property is not verified, we can confidently reject the $(A,\phi)$-modelling \autoref{assumpt:structure}. This enables us to consider a rich set of potential features $\left((A^1, \phi^1), \dots, (A^M, \phi^M)\right)$ rather than relying on a single assumption.

@@

\begin{assumption}[Multi-model ambiguity]
\label{assumpt:multi-model-ambiguity}
Assume that the true dynamics $f$ lies within a finite set of candidate models $f^1, \dots, f^M$.
\begin{equation}
\exists m\in[M]: \dot{x}(t) = f^M(x(t), u(t)), \forall t\geq 0
\end{equation}
\end{assumption}
We want to adapt our planning algorithm in order to balance these concurrent hypotheses in a robust fashion, i.e. maximise a robust objective with discrete ambiguity:
\begin{equation}
\label{eq:robust-objective-discrete}
V^r = \sup_{\bu}\min_{m\in[M]} \sum_{t=1}^\infty \gamma^t R_t^m
\end{equation}
where $R_t^m$ is the reward obtain by following the control sequence $\bu$ up to step $t$ under the dynamics $f^m$.
This objective could be optimised in the same way as \ref{sec:control}, by computing a hull over the $M$ possible trajectories, but this would result in a coarse and lossy approximation. Instead, we can exploit the finite uncertainty structure of \autoref{assumpt:multi-model-ambiguity} to asymptotically recover the true $V^r$ by modifying the \texttt{OPD} planning algorithm in the following way:

@@

Note that this theorem bounds the performance gap with respect to the number of node expansions, but each expansion still requires $M$ time more simulations in the multi-model setting than in the single-model setting. The solution of the robust objective \eqref{eq:robust-objective-discrete} with discrete ambiguity $f\in\{f^m\}_{m\in[M]}$ can be recovered exactly, asymptotically when the planning budget $H$ goes to infinity. This is in contrast to the results obtained \autoref{sec:control} regarding the robust objective \eqref{eq:robust-control} with continuous ambiguity $A(\theta)\in\cC_{[N],\delta}$, for which there is a gap between the asymptotic performance and the true robust value, as discussed in \autoref{thm:control-error}.

@@

\paragraph{Obstacle avoidance with unknown friction}
We first consider a simple illustrative example: the control of a 2D system with position $(x,y)$ and velocity $(v_x, v_y)$ moving by means of a force $(u_x, u_y)$ in an environment with unknown anisotropic friction. The dynamics are:

\begin{equation*}
\begin{bmatrix}
\dot{x}\\
\dot{y}\\
\dot{v_x}\\
\dot{v_y}\\
\end{bmatrix} = 
\begin{bmatrix}
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1 \\
0 & 0 & -\theta_x & 0 \\
0 & 0 & 0 & -\theta_y
\end{bmatrix}
\begin{bmatrix}
{x}\\
{y}\\
{v_x}\\
{v_y}\\
\end{bmatrix}
+
\begin{bmatrix}
0\\
0\\
{u_x}\\
{u_y}\\
\end{bmatrix}
\end{equation*}

The reward is non-smooth and encodes the task of reaching a goal position $x_g$ while avoiding collisions: $R(x) = \delta(x)/(1 + \|x - x_g\|_2)$  where $\delta(x)$ is $0$ whenever $x$ overlaps with an obstacle position.

\begin{table}[htbp]
	\caption{Mean and worst-case performance}
	\label{tab:results}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		\multicolumn{1}{l}{Task} &
		\multicolumn{2}{c}{Obstacle}    &
		\multicolumn{2}{c}{Driving}\\
		\cmidrule(lr){2-3}
		\cmidrule(lr){4-5}
		Perf. &
		\multicolumn{1}{c}{mean} &
		\multicolumn{1}{c}{min} &
		\multicolumn{1}{c}{mean} &
		\multicolumn{1}{c}{min}\\
		\midrule
		Oracle & \num{0} & \num{0} & \num{0} & \num{0} \\
		Nominal & \num{0} & \num{0} & \num{0} & \num{0} \\
		\OurAlgorithm & \num{0} & \num{0} & \num{0} & \num{0} \\
		\bottomrule
	\end{tabular}
\end{table}

In  \autoref{fig:regret} study the evolution of the agent conservativeness as they collect more data. Given a trajectory $x_0, R_0, x_1, R_1, \dots$, we compute the cumulative return $J_n(x_n) = \sum_{k\geq n} \gamma^k R_k$ at each state, and compare it to the optimal value $V(x_n)$, the return the agent would have got if it had acted optimally with respect to the oracle model.

\begin{figure}[tp]
	\centering
	\includegraphics[width=0.8\linewidth]{img/regret.pdf}
	\caption{We compare the performance of the robust agent to the optimal performance $\hat{V}^r(s_n)-V^*(s_n)$, with respect to the number $n$ of interactions.}
	\label{fig:regret}
\end{figure}

@@

\section*{Conclusion}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum

@@

\subsection{Proof of \autoref{thm:control-error}}

We first bound the model estimation error.
\begin{lemma}
Under sufficient excitations in the features 
\begin{align}
\label{eq:excitation}
\exists \underline{\phi},\overline{\phi}>0, n_0: \forall n\geq n_0,i,j\in[p],\nonumber\\ \underline{\phi}^2 \leq \sum_k \sum_l \phi_{n,k,i}\Sigma_{p,k,l}^{-1}\phi_{n,l,j} \leq \overline{\phi}^2,
\end{align}
then,
\[\|\theta - \theta_{[N],\lambda}\|_{2} = \cO\left( \sqrt{\frac{\log N}{N}} \right) \]
\end{lemma}
\begin{proof}
By \eqref{eq:g_n_lambda} and \eqref{eq:excitation}, we have $G_{[N],\lambda,i,j} \geq N\underline{\phi}^2 $

Hence, by \eqref{eq:confidence-ellipsoid} we have 
\begin{align*}
	\|\theta - \theta_{[N],\lambda}\|_{G_{[N],\lambda}} \geq \sqrt{N}\underline{\phi} \|\theta - \theta_{[N],\lambda}\|_{2} 
\end{align*}
and \eqref{eq:beta_n} gives

\begin{align*}
\beta_N(\delta) &= \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
+ (\lambda d)^{1/2}S\\
&\leq \sqrt{d\log (N\overline{\phi}^2) / \lambda^{d/2}} + \cO(1)
\end{align*}

which yields the desired bound.
\end{proof}

Then, we propagate this estimation error through the state prediction.

\begin{lemma}
	\todo{Which result?}
\end{lemma}
\begin{proof}
Let $\delta x = \ox - \ux$. Assuming that eventually after $n\geq n_0$ $A_0$ becomes close enough to $A(\theta)$ to always be Metlzer. Then, \eqref{eq:interval-predictor} gives the dynamics:
\begin{align*}
\dot{\delta x} = A_0\delta x + |\Delta A|(\ox^+ + \ux^-) + |D|(\overline{\omega} - \underline{\omega})
\end{align*}
where $|M| = M^+ + M^-$.

Moreover, it holds that $-\ux^+ -\ox^- \leq \delta x \leq \ox^+ + \ux^-$, which implies $|\ux^+ + \ox^-| \leq 2 |\delta x|$.
